Fractional Decomposition Tree Algorithm: A tool for studying the integrality gap of Integer Programs

### 2020-08-02 FDT2020/Introduction.tex
Cindys Intro comments replied

Many researchers (see \cite{davids,vazirani}) have developed polynomial time LP-based approximation algorithms that find solutions for special classes of IPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. If the analysis uses the LP bound to prove the approximation quality, then $C$ is at least the integrality gap.
%However, for many combinatorial optimization problems there is a limit to such techniques based on LP relaxations, represented  by the {integrality gap} of the IP formulation.
\cindy{Arash: Just read and comment it out if you're OK with this. I'm not sure I got an answer for the following questions, so I'll just make it clear this is the definition we use: Does the integrality gap always require positive objective coefficients?  Or is that an extra restriction in our definition?  Is there a reference that has that restriction?} 
\begin{definition}
The integrality gap $g(I)$ for instance $I$ is: $$g(I)= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)},$$
where $z_{IP}(I,c)$ is the optimal solution to the integer program and $z_{LP}(I,c)$ is the solution to the linear-programming relaxation.
\end{definition}

@@

We give a general approximation framework for solving binary IPs.
Consider the set of points described by sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume in addition that $S(I),P(I)\subseteq [0,1]^n$.
\cindy{Is it $S(I) \subseteq \{0,1\}^n$?}
For a vector $x\in \mathbb{R}_{\geq 0}^n$ such that $x\in P(I)$, let the {\em support} of $x$ be $\spp(x)= \{i \in [n]: x_i \neq 0\}$. For an integer $\beta$ let $\{\beta\}^n$ be the vector $y\in \mathbb{R}^n$ with $y_i=\beta$ for $i\in [n]$.

@@

\cindy{Arash, please check how I defined $L(I)$.  It was not defined before.}
\cindy{TODO: This implies we know $g(I)$, which is not true in general.  We should add how we bound/estimate it.  I'll return to this after putting that detail back in core if you (Arash) haven't added it by then.}
Here is how the FDT algorithm works at a high level for an instance $I$ with LP feasible region $L(I)$: in iteration $i$ the algorithm maintains a convex combination of  vectors in $\mathcal{D}(L(I))$ that have a 0 or 1 value for coordinates indexed $0,\ldots,i-1$. Let $y$ be a vector in the convex combination from iteration $i-1$. We solve a linear programming problem that gives us $\theta\in [0,1]$ and $y^0,y^1\in \mathcal{D}(L(I))$ such that $g(I) y\geq \theta y^0 + (1-\theta) y^1$ and $y^0_i=0$ and $y^1_i=1$. We then replace $y$ in the convex combination with $\frac{\theta}{g(I)}y^0 +\frac{1-\theta}{g(I)}y^1$. Repeating this for every vector in the convex combination from the previous iteration yields a new convex combination of points. It is ``more'' integral because now all vectors in the convex combination are integral in their first $i+1$ elements. If in any iteration there are too many points in the convex combination,  we solve a linear programming problem that ``prunes'' the convex combination. At the end we have a convex combination of integer solutions $\mathcal{D}(L(I))$. For each such solution $z$ we invoke the DomToIP algorithm (see Section \ref{domTOIP}) to find $z'\in S(I)$ where $z'\leq z$.

>>>>

Many researchers (see \cite{davids,vazirani}) have developed polynomial time LP-based approximation algorithms that find solutions for special classes of IPs whose cost are provably smaller than $C\cdot z_{LP}(I,c)$. If the analysis uses the LP bound to prove the approximation quality, then $C$ is at least the integrality gap.
%However, for many combinatorial optimization problems there is a limit to such techniques based on LP relaxations, represented  by the {integrality gap} of the IP formulation.
\cindy{Arash: Just read and comment it out if you're OK with this. I'm not sure I got an answer for the following questions, so I'll just make it clear this is the definition we use: Does the integrality gap always require positive objective coefficients?  Or is that an extra restriction in our definition?  Is there a reference that has that restriction?} 
\arash{I looked for a formal definition of integrality gap in the literature this is the most descriptive I've found from \cite{davids} as follows: \textit{the integrality gap of an integer program is the worst-case ratio over all instances of the problem of value of an optimal solution to the integer programming formulalation to value of an optimal solution to its linear programming relaxation.}. Based on this definition I think we are not adding a restriction but we are looking at specific type of IPs arising form problems whose instances have non-negative cost vectors which is not a huge stretch as most combinatorial optimization problems are of this forms.}
\begin{definition}
The integrality gap $g(I)$ for instance $I$ is: $$g(I)= \max_{c\geq 0}\frac{z_{IP}(I,c)}{z_{LP}(I,c)},$$
where $z_{IP}(I,c)$ is the optimal solution to the integer program and $z_{LP}(I,c)$ is the solution to the linear-programming relaxation.
\end{definition}

@@

We give a general approximation framework for solving binary IPs.
Consider the set of points described by sets $S(I)$ and $P(I)$ as in (\ref{S}) and (\ref{P}), respectively. Assume in addition that $S(I)\in \{0,1\}^n$ and $P(I)\subseteq [0,1]^n$.
%\cindy{Is it $S(I) \subseteq \{0,1\}^n$?}
For a vector $x\in \mathbb{R}_{\geq 0}^n$ such that $x\in P(I)$, let the {\em support} of $x$ be $\spp(x)= \{i \in [n]: x_i \neq 0\}$. For an integer $\beta$ let $\{\beta\}^n$ be the vector $y\in \mathbb{R}^n$ with $y_i=\beta$ for $i\in [n]$.

@@

\cindy{Arash, please check how I defined $L(I)$.  It was not defined before.}
\cindy{TODO: This implies we know $g(I)$, which is not true in general.  We should add how we bound/estimate it.  I'll return to this after putting that detail back in core if you (Arash) haven't added it by then.}
\arash{Isn't $P(I)$ the feasible region defined above. I keep $L(I)$ for now but we can replace it later if needed.} \arash{I fixed below so that I don't imply we know $g(I)$, we also might to clearly state that somewhere?}

Here is how the FDT algorithm works at a high level for an instance $I$ with LP feasible region $L(I)$: in iteration $i$ the algorithm maintains a convex combination of  vectors in $\mathcal{D}(L(I))$ that have a 0 or 1 value for coordinates indexed $0,\ldots,i-1$. Let $y$ be a vector in the convex combination from iteration $i-1$. We solve a linear programming problem that gives us $\theta_0,\theta_1\in [0,1]$ and $y^0,y^1\in \mathcal{D}(L(I))$ such that $y\geq \theta_0 y^0 + \theta^1 y^1$, $\theta_0+\theta_1\leq g(I)$, $y^0_i=0$ and $y^1_i=1$. We then replace $y$ in the convex combination with $\frac{\theta_0}{\theta_0+\theta_1}y^0 +\frac{\theta_1}{\theta_0+\theta_1}y^1$.  Repeating this for every vector in the convex combination from the previous iteration yields a new convex combination of points. It is ``more'' integral because now all vectors in the convex combination are integral in their first $i+1$ elements. If in any iteration there are too many points in the convex combination,  we solve a linear programming problem that ``prunes'' the convex combination. At the end we have a convex combination of integer solutions $\mathcal{D}(L(I))$. For each such solution $z$ we invoke the DomToIP algorithm (see Section \ref{domTOIP}) to find $z'\in S(I)$ where $z'\leq z$.
